---
title: "CourseraAssignment02"
output: html_document
date: "2024-01-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Introduction

Reference the [tidyverse style guide](https://style.tidyverse.org/) while completing the exercises. 

Review Sections 2.1-2.5 from Hands-On Programming in R and Sections 3.1 - 3.4 from Hands-On Programming in R and complete the following exercises. These exercises are ungraded.


#### Question 1. 

Write a roll function that simulates rolling two six-sided die. The function should contain no input arguments.

```{r roll}
roll <- function(){
  dice1<-sample(1:6,2,replace=TRUE)
  return(dice1)
}
result=roll()
print(result)
```

#### Question 2.

Rewrite the roll function to include an argument for the type of die (e.g. three-, six-, eight-sided). Think about what other input arguments should be allowed. How would you update your function to include them? 

```{r roll}
roll <- function(noOfDice,sides){
  dice1<-sample(1:sides,noOfDice,replace=TRUE)
  return(dice1)
}
result=roll(6,8)
print(result)
```

#### Question 3. 

Produce a historgram of 50,000 rolls of three eight-sided fair dice. 

```{r roll}
roll <- function(){
  
  dice1<-sample(1:8,3,replace=TRUE)
  return(sum(dice1))

}

num_rolls <- 50000
dice_results <- replicate(num_rolls, roll())

hist(dice_results, breaks = seq(2.5, 25.5, by = 1), col = "skyblue", border = "black", xlab = "Sum of Dice", ylab = "Frequency", main = "Histogram of Three Eight-sided Dice Rolls")

```
#### Question 4. 

Produce a historgram of 50,000 rolls of three eight-sided dice where the dice are loaded so that the number 7 has a higher probability of being rolled than the other numbers, assume all the other numbers have a 1/10 probability of being rolled.

```{r roll}
roll <- function(noOfDice,sides){
  probabilities <- c(1/10, 1/10, 1/10, 1/10, 1/10, 1/10, 3/10, 1/10)
  dice1<-sample(1:8,3,replace=TRUE,prob = probabilities)
  return(sum(dice1))
}

num_rolls <- 50000
dice_results <- replicate(num_rolls, roll())

hist(dice_results, breaks = seq(2.5, 25.5, by = 1), col = "skyblue", border = "black", xlab = "Sum of Dice", ylab = "Frequency", main = "Histogram of Three Eight-sided Dice Rolls")


```

#### Question 5. 

Rewrite the "rescale01()" function such that -Inf is mapped to 0 and Inf is mapped to 1.

```{r rescale}
rescale01 <- function(x) {
  # Replace -Inf with 0
  x[is.infinite(x) & x < 0] <- 0
  # Replace Inf with 1
  x[is.infinite(x) & x > 0] <- 1
  # Rescale to [0, 1]
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
}

values <- c(-Inf, 5, 10, Inf)
rescaled_values <- rescale01(values)
print(rescaled_values)



```


#### Question 6.

Write both_na(), a function that takes two vectors of the same length and returns the number of positions that have an NA in both vectors.

```{r na}

both_na <- function(vec1, vec2) {
  if (length(vec1) != length(vec2)) {
    stop("Vectors must be of the same length.")
  }

  na_positions <- sum(is.na(vec1) & is.na(vec2))
  
  return(na_positions)
}

vector1 <- c(1, NA, 3, NA, 5)
vector2 <- c(NA, 2, 3, NA, 5)

result <- both_na(vector1, vector2)
print(result)




```



