---
title: "Lab6-dplyr"
output: html_document
date: "2024-01-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Question 1. 

Use the following definitions of cancellation codes to add a "Code" column to hflights, then use tail(hflights$Code[!is.na(hflights$Code)], n = 20) to see that your replacement worked by looking at the last 20 entries that are not NA. 

"A"   =    "carrier"
"B"  =    "weather"
"C"   =   "FF"
"D"   =   "security"
"E"   =   "not cancelled"

```{r}
library(dplyr)
library(hflights)

# Assuming 'hflights' is your dataset

# Define cancellation codes
cancellation_codes <- c("A" = "carrier", "B" = "weather", "C" = "FF", "D" = "security", "E" = "not cancelled")

# Add 'Code' column based on cancellation codes
hflights$Code <- cancellation_codes[hflights$CancellationCode]

# Display the last 20 entries with non-NA 'Code'
tail(hflights$Code[!is.na(hflights$Code)], n = 20)


```

#### Question 2. 

What does the any_of() function do? Why might it be helpful in conjunction with this vector when analyzing nycflights13 data?

`vars <- c("year", "month", "day", "dep_delay", "arr_delay")`

```{r}
library(dplyr)
library(hflights)

vars <- c("year", "month", "day", "dep_delay", "arr_delay")
# Assuming df is your dataset
selected_columns <-  hflights %>% select(any_of(vars))
selected_columns



```

#### Question 3. 

Does the result of running the following code surprise you? How does the select helpers deal with case by default? How can you change that default?

`select(flights, contains("TIME"))`

```{r}
select(hflights, contains("TIME", ignore.case = TRUE))

```

#### Question 4. 

Use select to show only the columns of hflights that have the word time or delay in the variable name or have the carrier id.

```{r}
library(dplyr)

selected_columns <- hflights %>%
  select(contains("time", ignore.case = TRUE) | contains("delay", ignore.case = TRUE) | matches("carrier id", ignore.case = TRUE))

head(selected_columns)

```

#### Question 5. 

Find all flights from nycflights13 that had an arrival delay of two or more hours. Put the arrival delay first in your output so you can check it.

```{r}
library(dplyr)

delayed_flights <- nycflights13::flights %>%
  filter(arr_delay >= 120) %>%
  arrange(arr_delay) %>%
  select(arr_delay, everything())

head(delayed_flights)



```

#### Question 6. 

Find all flights from nycflights13 that were operated by United, American, or Delta.

```{r}

library(dplyr)

selected_flights <- nycflights13::flights %>%
  filter(carrier %in% c("UA", "AA", "DL"))

head(selected_flights)


```

#### Question 7.

Find all flights from nycflights13 that arrived more than two hours late.

```{r}
library(dplyr)

late_flights <- nycflights13::flights %>%
  filter(arr_delay > 120)

head(late_flights)


```

#### Question 8. 

Find all flights from nycflights13 that had a departure delay by at least an hour

```{r}
library(dplyr)

delayed_departures <- nycflights13::flights %>%
  filter(dep_delay >= 60)

head(delayed_departures)


```

#### Question 9. 

Sort flights from nycflights13 to find the 5 most flights whose arrival was delayed the longest. Find the 5 flights that left earliest (most ahead of scheduled departure time). 
```{r}
library(dplyr)

# Top 5 flights with the longest arrival delays
top_delayed_flights <- nycflights13::flights %>%
  arrange(desc(arr_delay)) %>%
  head(5)

# Top 5 flights that left earliest
earliest_departures <- nycflights13::flights %>%
  arrange(dep_time) %>%
  head(5)

# Display the results
print(top_delayed_flights)


print(earliest_departures)


```