---
title: "Lab6-Grouping"
output: html_document
date: "2024-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

### Introduction

Reference the [tidyverse style guide](https://style.tidyverse.org/) while completing the exercises. 

These exercises are ungraded.

#### Question 1. 

Using the nycflights13 data, calculate a new variable called "hr_delay" and arrange the flights dataset in order of the arrival delays in hours (longest delays at the top). Put the new variable you created just before the departure time. Hint: use the experimental argument .before. 

```{r}


library(nycflights13)
library(dplyr)

data("flights")


flights <- flights %>%
  mutate(hr_delay = arr_delay / 60) %>%
  arrange(desc(hr_delay))


flights <- flights %>%
  select(hr_delay, everything())


head(flights)



```
#### Question 2. 

What proportion of flights in nycflights13 are delayed by more than on hour? 

```{r}


library(nycflights13)
library(dplyr)


data("flights")

proportion_delayed <- flights %>%
  mutate(delayed = ifelse(arr_delay > 60, 1, 0)) %>%
  summarise(proportion = mean(delayed, na.rm = TRUE))


proportion_delayed$proportion




```


